# Explorative Data Analysis (EDA)

### Import libraries
```{r message=FALSE, warning=FALSE}

library(ggplot2)
library(ggcorrplot)
library(cowplot)
library(PerformanceAnalytics)
library(GGally)

```

### Import final and final_individual tables
```{r}

final <- read.csv("/Users/elypa/Downloads/final.csv")
final_individual <- read.csv("/Users/elypa/Downloads/final_individual.csv")

```

### Tidy up and remove unuseful columns ("id" and "area5")
```{r}

final <- final[-c(1, 10)]
final <- na.omit(final)

final_individual <- final[-c(1, 10)]
final_individual <- na.omit(final_individual)

```

### Compute variables correlation matrix

```{r message=FALSE, warning=FALSE}

cor(final)
cor(final_individual)

```

### Plot correlation matrix
```{r fig.height=9, fig.width=12, message=FALSE, warning=FALSE}

# final
# Correlation matrix
corr_final = round(cor(final), 1)

# Plot
final_corplot <- ggcorrplot(corr_final, hc.order = TRUE, 
                            type = "lower", 
                            lab = TRUE, 
                            lab_size = 3, 
                            method="circle", 
                            colors = c("tomato2", "white", "springgreen3"), 
                            title="final variable corelation", 
                            ggtheme=theme_bw)

#final_individual
# Correlation matrix
corr_final_indiv = round(cor(final_individual), 1)

# Plot
final_indv_corplot <- ggcorrplot(corr_final_indiv, hc.order = TRUE, 
                                 type = "lower", 
                                 lab = TRUE, 
                                 lab_size = 3, 
                                 method="circle", 
                                 colors = c("tomato2", "white", "springgreen3"), 
                                 title="final_individual variable corelation", 
                                 ggtheme=theme_bw)

plot_grid(final_corplot, final_indv_corplot)

```

#### As the correlographs show, the Exploratory Data Analysis (EDA) detected several high Pearson correlation coefficients. In particular, a $\rho > |\pm 0.6|$ is detected between:
#### **staciv-anasc**, **staciv-sex**, **anasc-ireg**, and **ireg-sex** in both datasets. A strong negative correlation ($\rho = -1$) characterises the couple **sex-anasc**.
#### These results can already cut out some classification methods having the assumption of independence between predictors.

### Compute more informative matrices
```{r fig.height=16, fig.width=16, message=FALSE, warning=FALSE}

final_chart <- ggpairs(final,            
                       lower = list(continuous = wrap("points", alpha = 0.5)), 
                       diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 ))
                       )

final_chart


```

```{r fig.height=16, fig.width=16, message=FALSE, warning=FALSE}

final_indiv_chart <- ggpairs(final_individual,            
                       lower = list(continuous = wrap("points", alpha = 0.5)), 
                       diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 ))
                       )

final_indiv_chart

```

